\section{Extended Problem Examples}
\label{app:examples}

We present six complete problem--solution pairs from \ecagbench{}, selected to illustrate the range of problem types, difficulty levels, and mathematical content.

\subsection*{A.1: Counterexample --- Infinite Disjoint Union (Level 2)}

\begin{problembox}[title={\ecagid{0024} --- Basic Concepts}]
\textbf{Problem:} Let $k$ be a field. Is the scheme $\coprod_{i=1}^{\infty} \Spec(k)$ affine? Justify your answer by identifying which property of affine schemes it fails to satisfy.

\medskip
\textbf{Solution (abbreviated):}
The scheme $X = \coprod_{i=1}^{\infty} \Spec(k)$ is not affine.
The key property that fails is \emph{quasi-compactness}: affine schemes are quasi-compact, but $X$ has the discrete topology on an infinite set, so the cover $\{U_i = \Spec(k)\}$ has no finite subcover.

Alternatively, if $X$ were affine, we would have $X \cong \Spec(R)$ for $R = \Gamma(X, \mathcal{O}_X) \cong \prod_{i=1}^{\infty} k$.
But $\Spec(\prod_{i=1}^{\infty} k)$ is a connected scheme (it has no non-trivial idempotents besides those coming from finite products), contradicting the fact that $X$ is disconnected with infinitely many connected components.

\medskip
\textbf{Type:} Counterexample \quad \textbf{Difficulty:} 2 \quad \textbf{Verification:} CAS (check quasi-compactness of global sections ring)
\end{problembox}

\subsection*{A.2: Example --- Hopf Algebra Structures in Char $p$ (Level 1)}

\begin{problembox}[title={\ecagid{0003} --- Basic Concepts}]
\textbf{Problem:} Let $k$ be a field of characteristic $p > 0$. Write down explicitly the Hopf algebra structures (comultiplication, counit, antipode) on $k[x]/(x^p - 1)$ for $\mu_p$ and on $k[x]/(x^p)$ for $\alpha_p$. Using the substitution $u = x - 1$ to identify the underlying algebras as $k[u]/(u^p)$, show that $\mu_p \not\cong \alpha_p$ as group schemes.

\medskip
\textbf{Solution (abbreviated):}
Under $u = x-1$, both algebras become $k[u]/(u^p)$, but the comultiplications differ:
\begin{align*}
\Delta_\mu(u) &= u \otimes 1 + 1 \otimes u + u \otimes u \quad (\text{from } \mu_p) \\
\Delta_\alpha(u) &= u \otimes 1 + 1 \otimes u \quad (\text{from } \alpha_p)
\end{align*}
Any $k$-algebra automorphism of $k[u]/(u^p)$ has the form $u \mapsto \lambda u + (\text{higher terms})$ with $\lambda \neq 0$, but $\Delta_\mu$ has a quadratic term $u \otimes u$ that cannot be eliminated by any such automorphism.

\medskip
\textbf{Type:} Example \quad \textbf{Difficulty:} 1 \quad \textbf{Verification:} CAS (verify Hopf algebra axioms)
\end{problembox}

\subsection*{A.3: Counterexample --- Locally Finite Type vs.\ Finite Type (Level 2)}

\begin{problembox}[title={\ecagid{0043} --- Basic Concepts}]
\textbf{Problem:} State the precise difference between ``locally of finite type'' and ``of finite type'' for a morphism of schemes. Give an example of a morphism that is locally of finite type but not of finite type, and identify which condition fails.

\medskip
\textbf{Solution (abbreviated):}
A morphism $f: X \to Y$ is \emph{locally of finite type} if for every $x \in X$, there exist affine neighborhoods $U = \Spec(A)$ of $f(x)$ and $V = \Spec(B) \subseteq f^{-1}(U)$ of $x$ such that $B$ is a finitely generated $A$-algebra. It is \emph{of finite type} if additionally $f$ is quasi-compact.

The morphism $\coprod_{i \in I} \mathbf{A}^1_k \to \mathbf{A}^1_k$ (identity on each component, $I$ infinite) is locally of finite type but not of finite type, because the source is not quasi-compact.

\medskip
\textbf{Type:} Counterexample \quad \textbf{Difficulty:} 2 \quad \textbf{Verification:} CAS
\end{problembox}

\subsection*{A.4: Example --- \'Etale Cover of Nodal Cubic (Level 2)}

\begin{problembox}[title={\ecagid{0158} --- Basic Concepts}]
\textbf{Problem:} Let $k$ be an algebraically closed field with $\chr(k) \neq 2$, and let $Y = \Spec(k[x,y]/(y^2 - x^2(x+1)))$ be the nodal cubic. Construct a degree-2 \'etale morphism $f: X \to Y$ and compute the fibers over the node and a smooth point.

\medskip
\textbf{Solution (abbreviated):}
Define $X = \Spec(k[s,t]/(t^2 - (s^2-1)^2))$ with $f$ given by $x \mapsto s^2 - 1$, $y \mapsto st$.
The fiber over the node $(0,0)$ is $\Spec(k[s,t]/(s^2-1, t^2-(s^2-1)^2, st)) = \{(1,0), (-1,0)\}$, a reduced scheme of two points.
At smooth points, $f$ is \'etale of degree~2 (one checks that $\Omega_{X/Y} = 0$ away from the node).

\medskip
\textbf{Type:} Example \quad \textbf{Difficulty:} 2 \quad \textbf{Verification:} CAS (verify \'etale condition, compute fibers)
\end{problembox}

\subsection*{A.5: Computation --- Horrocks--Mumford Bundle (Level 2)}

\begin{problembox}[title={\ecagid{0230} --- Computations}]
\textbf{Problem:} Let $\mathscr{F}$ be the Horrocks--Mumford bundle on $\mathbf{P}^4$ with $c_1(\mathscr{F}) = 5h$ and $c_2(\mathscr{F}) = 10h^2$. Using HRR, compute $\chi(\mathscr{F}(n-5))$ as a closed-form polynomial in $n$.

\medskip
\textbf{Solution (abbreviated):}
The Chern character of $\mathscr{F}(m)$ (where $m = n-5$) is computed from $c_1$ and $c_2$ using the splitting principle, then paired with the Todd class of $\mathbf{P}^4$.
The result is a degree-4 polynomial in $n$, which at $n=5$ yields $\chi(\mathscr{F}) = h^0 - h^1 = 4 - 2 = 2$, consistent with known cohomology.

\medskip
\textbf{Type:} Computation \quad \textbf{Difficulty:} 2 \quad \textbf{Verification:} Exact match + CAS
\end{problembox}

\subsection*{A.6: Example --- Proper Birational to Normal (Level 2)}

\begin{problembox}[title={\ecagid{0019} --- Basic Concepts}]
\textbf{Problem:} Let $f: X \to Y$ be a proper birational morphism with $Y$ normal. Prove that $f_* \mathcal{O}_X \cong \mathcal{O}_Y$. Which property of normal schemes is used?

\medskip
\textbf{Solution (abbreviated):}
The structure map $\mathcal{O}_Y \to f_* \mathcal{O}_X$ is injective (since $f$ is birational and dominant).
For surjectivity, any section $s \in (f_* \mathcal{O}_X)(U) = \mathcal{O}_X(f^{-1}(U))$ restricts to a rational function on $U$ that is regular on the dense open $f^{-1}(U \cap V)$ where $f$ is an isomorphism.
Since $Y$ is normal, the algebraic Hartogs lemma (Serre's $R_1 + S_2$ criterion) extends $s$ to a regular function on $U$.

\medskip
\textbf{Type:} Example \quad \textbf{Difficulty:} 2 \quad \textbf{Verification:} Rubric
\end{problembox}

\section{Evaluation Prompt and Scoring Rubric}
\label{app:prompt}

\subsection*{B.1: Full Evaluation Prompt}

\begin{lstlisting}
You are an expert algebraic geometer with deep
knowledge of scheme theory, sheaf cohomology,
moduli spaces, and computational algebraic geometry.

PROBLEM:
{problem}

INSTRUCTIONS:
Provide a CONCRETE construction with full
mathematical justification. Your response should
include:

1. **The explicit mathematical object**: State the
   object clearly (e.g., a specific polynomial,
   ring, scheme, morphism, sheaf, or cohomology
   class).

2. **Verification of all required properties**: For
   each property mentioned in the problem, provide
   a rigorous argument that your object satisfies
   (or, for counterexamples, violates) it.

3. **Key insight or theorem used**: Identify the main
   mathematical idea underlying your construction.

FORMAT:
- State your answer at the very beginning, clearly
  labeled.
- Then provide the detailed justification.
- Use standard LaTeX notation for mathematical
  expressions.
- If the problem asks for a counterexample,
  explicitly state which hypothesis fails and why.
\end{lstlisting}

\subsection*{B.2: Scoring Rubric}

\begin{description}[nosep]
  \item[Score 1.0 (Correct):] \hfill
  \begin{itemize}[nosep]
    \item (C1) The constructed object is well-defined and belongs to the correct mathematical category.
    \item (C2) All properties stated in the problem are rigorously verified (or, for counterexamples, the target property is rigorously shown to fail).
    \item (C3) The answer is non-trivial and non-degenerate.
    \item (C4) The mathematical reasoning is sound and references to standard results are accurate.
  \end{itemize}
  \item[Score 0.5 (Partial):] \hfill
  \begin{itemize}[nosep]
    \item The object is valid (C1 satisfied) but one or more of C2--C4 is partially satisfied.
    \item Typical cases: correct object with incomplete verification, or correct approach with a computational error that does not invalidate the construction.
  \end{itemize}
  \item[Score 0.0 (Incorrect):] \hfill
  \begin{itemize}[nosep]
    \item The object is invalid (C1 fails), or the properties are not satisfied (C2 fails fundamentally).
    \item Includes fabricated objects (F1), reversed properties (F2), and reliance on non-existent theorems (F5).
  \end{itemize}
\end{description}

\section{Dataset Schema and Access}
\label{app:schema}

\subsection*{C.1: JSON Schema}

Each entry in the \ecagbench{} dataset (\texttt{benchmark/data/ecag\_bench.jsonl}) conforms to the following schema:

\begin{lstlisting}
{
  "id": "ecag-XXXX",
  "problem": "Problem statement in LaTeX",
  "solution": "Full solution text",
  "answer": "Closed-form answer or null",
  "answer_type": "example|counterexample|
                  closed_form|proof|true_false",
  "difficulty": 1|2|3,
  "chapter": "basic_concepts|cohomology|
              computations|moduli|relative",
  "topic": "...",
  "subtopic": "...",
  "question_type": "example|counterexample|
                    computation|proof|true_false",
  "tags": ["..."],
  "source": "Original|Hartshorne X.Y.Z|...",
  "prerequisites": ["..."],
  "msc_codes": ["14XXX"],
  "verification_strategy": "exact|cas|rubric|hybrid",
  "verification_script": "path or null",
  "accepts_multiple": true|false,
  "is_stub": false,
  "latex_source": "path to .tex file"
}
\end{lstlisting}

Difficulty levels: 1 = routine (follows from definitions), 2 = non-trivial (requires significant theorem), 3 = deep (combines multiple areas or requires specialized knowledge).

\subsection*{C.2: Access}

\begin{itemize}[nosep]
  \item \textbf{GitHub:} {\small\url{https://github.com/Waerden001/Examples_and_Counterexamples_in_Elementary_Algebraic_Geometry}}
  \item \textbf{Website:} The accompanying textbook is available as a searchable website via GitHub Pages.
  \item \textbf{License:} CC-BY-4.0
\end{itemize}

\section{Case Study Full Outputs}
\label{app:case-studies}

This appendix provides representative model outputs from the pilot evaluation discussed in Section~\ref{sec:experiments}.
Outputs are from Claude Opus~4.6 unless otherwise noted.
The complete raw outputs for all 100 problems---including both Opus and Sonnet responses with associated scores and failure mode annotations---are released as JSONL files in the project repository (\texttt{eval/results/}), enabling independent analysis by other researchers.

\subsection*{D.1: Successful Level~3 Construction (\ecagid{0363})}

\textbf{Problem (verbatim from JSONL):} Using genus-0 Gromov--Witten invariants and Schubert calculus on $\operatorname{Gr}(2, n+1)$, compute $\langle H^i \mid H^j \mid H^k \rangle_{H^{n-1}}$ for $\mathbb{P}^n$ and derive the quantum cohomology ring $QH^*(\mathbb{P}^n, \mathbb{Z})$.

\textbf{Type:} Example \quad \textbf{Difficulty:} Level~3 \quad \textbf{Opus Score:} 1.0 \quad \textbf{Sonnet:} Timeout

\textbf{Model output (Opus, verbatim excerpt from the 5{,}443-character response; Markdown headings and display math adapted to \LaTeX\ formatting):}

\begin{quote}
\small
\textbf{Step 2 (continued). Evaluation maps and tautological classes}

The three evaluation maps $\operatorname{ev}_a: \overline{M}_{0,3}(\mathbb{P}^n,1) \to \mathbb{P}^n$ project onto the $a$-th copy of $\mathcal{U}$, then compose with $\operatorname{ev}: \mathcal{U} \to \mathbb{P}^n$. Set
$$\xi = c_1(\mathcal{O}_{\mathbb{P}(\mathcal{S})}(1)) \in H^2(\mathcal{U}).$$
Since $\operatorname{ev}^*(H) = \xi$ (the tautological class on the projectivized bundle equals the pullback of the hyperplane class), we get $\operatorname{ev}_a^*(H^{a_i}) = \xi_a^{a_i}$, where $\xi_a$ denotes $\xi$ on the $a$-th fiber factor.

\textbf{Step 3. Fiber integration via Segre classes}

The projection $\rho: \mathcal{U}^3_G \to G$ has fiber $(\mathbb{P}^1)^3$. Since $\xi_a$ lives on independent fiber factors, fiber integration gives
$$\rho_*(\xi_1^{a_1}\cdot \xi_2^{a_2}\cdot \xi_3^{a_3}) = \pi_*(\xi^{a_1})\cdot \pi_*(\xi^{a_2})\cdot \pi_*(\xi^{a_3}).$$
By the projective bundle formula for the rank-2 bundle $\mathcal{S}$:
$$\pi_*(\xi^a) = s_{a-1}(\mathcal{S})$$
where $s_j(\mathcal{S}) = c_j(\mathcal{S})^{-1}$ is the $j$-th Segre class. Since $c(\mathcal{S})\cdot c(\mathcal{Q}) = 1$ (with $\mathcal{Q}$ the quotient bundle), we have $s(\mathcal{S}) = c(\mathcal{Q})$, and for $\operatorname{Gr}(2,n+1)$ the Chern classes of $\mathcal{Q}$ are the special Schubert classes:
$$s_j(\mathcal{S}) = c_j(\mathcal{Q}) = \sigma_j, \qquad 0 \le j \le n-1.$$

\textbf{Step 4. Reduction to Schubert calculus}

Combining Steps 2--3:
$$\langle H^i, H^j, H^k\rangle_1 = \int_G \rho_*(\xi_1^i\cdot\xi_2^j\cdot\xi_3^k) = \int_{\operatorname{Gr}(2,\,n+1)} \sigma_{i-1}\cdot\sigma_{j-1}\cdot\sigma_{k-1}.$$

[\ldots Steps 5--6 apply Pieri's rule to evaluate the Schubert intersection, showing each admissible degree-1 invariant equals~1\ldots]

\textbf{Step 7. The quantum cohomology ring}

$$QH^*(\mathbb{P}^n, \mathbb{Z}) \;\cong\; \mathbb{Z}[H, q]\,\big/\,(H^{n+1} - q).$$

As a $\mathbb{Z}[q]$-module, this is free of rank $n+1$ with basis $\{1, H, H^2, \ldots, H^n\}$, deforming the classical relation $H^{n+1} = 0$ to $H^{n+1} = q$.

\textbf{Key insight}

The critical bridge between Gromov--Witten theory and Schubert calculus is the identification of $\overline{M}_{0,3}(\mathbb{P}^n, 1)$ as the triple fiber product $\mathbb{P}(\mathcal{S})^3_G$ over the Grassmannian of lines. Fiber integration converts evaluation-map pullbacks $\operatorname{ev}_a^*(H^{a_i}) = \xi^{a_i}$ into Segre classes $\pi_*(\xi^{a_i}) = \sigma_{a_i - 1}$, reducing the enumerative GW computation to a classical Schubert intersection. The rigidity of lines (uniqueness of $\sigma_{(n-1,n-1)}$ as the point class) then forces every admissible degree-1 invariant to equal 1, yielding the single deformation $H^{n+1} = q$.
\end{quote}

This Level~3 problem requires synthesizing Gromov--Witten theory, Schubert calculus, and quantum cohomology.
Opus produced a correct 5{,}443-character response deriving $QH^*(\mathbb{P}^n, \mathbb{Z}) \cong \mathbb{Z}[H, q]/(H^{n+1} - q)$, while Sonnet timed out after 600 seconds---illustrating the capability gap at the highest difficulty level.

\subsection*{D.2: Property Confusion (\ecagid{0174}, F2)}

\textbf{Problem (verbatim from JSONL):} Let $R = k[x,y,z]/(xy - z^2)$ where $\operatorname{char}(k) \neq 2$, and let $\mathfrak{m} = (x,y,z)$. Prove that $\widetilde{\mathfrak{m}}$ is a reflexive sheaf on $\operatorname{Spec}(R)$ that is not locally free by showing $\mathfrak{m}$ corresponds to a Weil divisor that is not Cartier, and that $\operatorname{Cl}(R) \cong \mathbb{Z}/2\mathbb{Z}$.

\textbf{Type:} Counterexample \quad \textbf{Difficulty:} Level~2 \quad \textbf{Score:} 0.5 \quad \textbf{Failure mode:} F2

\textbf{Model output (Opus, verbatim excerpt of the ``Answer'' section; bold markers from Markdown adapted to \LaTeX):}

\begin{quote}
\small
The sheaf $\widetilde{\mathfrak{m}}$ on $X = \operatorname{Spec}(R)$, where $R = k[x,y,z]/(xy - z^2)$ and $\mathfrak{m} = (x,y,z)$, is a reflexive sheaf of rank 1 that is \textbf{not} locally free. It corresponds to the Weil divisor class of the ruling $D = V(x,z)$, which generates $\operatorname{Cl}(R) \cong \mathbb{Z}/2\mathbb{Z}$. The divisor $D$ is not Cartier (so $\widetilde{\mathfrak{m}}$ is not locally free), but $2D$ is Cartier (cut out by $x$), so $\widetilde{\mathfrak{m}}$ is reflexive. [\ldots]
\end{quote}

\textbf{Analysis:} The model correctly identified the quadric cone, computed $\operatorname{Cl}(R) \cong \mathbb{Z}/2\mathbb{Z}$ rigorously using Nagata's sequence, and identified $D = V(x,z)$ as a non-Cartier Weil divisor with $\operatorname{div}(x) = 2D$.
However, the model then correctly discovered that $\widetilde{\mathfrak{m}}$ is \emph{not} reflexive (since $\mathfrak{m}$ has grade $\geq 2$ in the normal ring $R$), contradicting the problem's stated claim.
The model pivoted to showing that $\mathfrak{p} = (x,z)$ gives the correct reflexive non-locally-free sheaf instead.
This exemplifies F2: the model addresses a modified version of the problem rather than the one posed, even though the mathematical reasoning is sophisticated and correct for the modified statement.

\subsection*{D.3: Cross-Model Comparison (\ecagid{0090})}

\textbf{Problem (verbatim from JSONL):} In Hironaka's construction, let $X = \mathbb{P}_k^3$ with two smooth conics $C, D$ meeting transversely at $P, Q$. Explain why the $1$-cycle $L_1 + M_1$ on the resulting variety $H$ is algebraically equivalent to $0$, and why this implies $H$ is not projective.

\textbf{Type:} Example \quad \textbf{Difficulty:} Level~3

\textbf{Opus:} Score 0.5 (F6, incomplete verification). The model correctly set up the two blow-up orderings and the gluing construction, and gave a correct non-projectivity argument (an ample divisor must pair positively with effective cycles, contradicting algebraic equivalence to~$0$). However, the derivation of $L_1 + M_1 \sim_{\mathrm{alg}} 0$ contained logical errors: the model's intermediate linear equivalence relations actually yield $L_1 + M_2 \sim 0$ rather than the required $L_1 + M_1 \sim 0$, and the model incorrectly deduced that $L_1 \sim 0$ and $M_1 \sim 0$ individually.

\textbf{Sonnet:} Timeout. The model failed to produce any response within 600 seconds.

This case illustrates a common Level~3 pattern: Opus can engage with the problem and produce partially correct responses, while Sonnet fails to respond at all.
The failure mode (F6, incomplete verification) is less severe than a timeout, as it provides a correct partial construction that could be completed with CAS feedback in a multi-turn setting.
