\section{Experiments}
\label{sec:experiments}

We report results from a pilot evaluation designed to establish initial baselines, validate the benchmark design, and investigate the example-versus-counterexample performance gap.
We evaluate two frontier models from the same family to enable both absolute measurement and cross-model comparison.

\subsection{Evaluation Setup}

\paragraph{Models.}
We evaluate two Claude models from Anthropic~\citep{anthropic2024claude}: \textbf{Claude Opus~4.6} (the flagship model) and \textbf{Claude Sonnet~4.6} (the mid-tier model), both accessed via the Claude Code CLI (\texttt{claude -p}) with default generation settings.
Evaluating two models from the same family isolates the effect of model scale while controlling for training methodology.

\paragraph{Subset selection.}
From the full 332-problem benchmark, we select a stratified random subset of 100 problems.
The stratification ensures proportional representation across chapter, difficulty level, and question type, with minimum constraints of 20 counterexamples, 20 Level~1, and 10 Level~3 problems.
The resulting subset contains 70 examples, 23 counterexamples, 4 computations, and 3 proofs, distributed as: 47 Basic Concepts, 21 Computations, 17 Moduli, 15 Cohomology; and 22 Level~1, 67 Level~2, 11 Level~3.

\paragraph{Prompt and scoring.}
Each problem is presented using the prompt template in Appendix~\ref{app:prompt}.
Responses were scored using a hybrid pipeline: an LLM judge evaluated each response against the human-authored reference solution on the 0/0.5/1 scale, followed by manual review of all partial-credit and sampled full-credit cases.
Problems where the model timed out (no response within 600 seconds) were scored as 0.0.
We discuss limitations of this scoring approach in Section~\ref{sec:self-eval-limitations}.

\subsection{Pilot Results}

\paragraph{Overall performance.}
Table~\ref{tab:pilot-results} presents the primary results for both models.

\begin{table}[ht]
\centering
\caption{Pilot evaluation results on 100-problem stratified subset. Accuracy is the proportion of scores $= 1.0$. Timeouts ($>$600s, no response) are counted as incorrect.}
\label{tab:pilot-results}
\begin{tabular}{lcccccc}
\toprule
& \textbf{$n$} & \textbf{Correct} & \textbf{Partial} & \textbf{Incorrect} & \textbf{Accuracy} & \textbf{Timeouts} \\
\midrule
\multicolumn{7}{l}{\textbf{Claude Opus 4.6}} \\
\quad Overall & 100 & 90 & 6 & 4 & 0.90 & 4 \\
\quad Level 1 & 22 & 21 & 1 & 0 & 0.95 & 0 \\
\quad Level 2 & 67 & 62 & 4 & 1 & 0.93 & 1 \\
\quad Level 3 & 11 & 7 & 1 & 3 & 0.64 & 3 \\
\midrule
\multicolumn{7}{l}{\textbf{Claude Sonnet 4.6}} \\
\quad Overall & 100 & 78 & 4 & 18 & 0.78 & 18 \\
\quad Level 1 & 22 & 21 & 0 & 1 & 0.95 & 1 \\
\quad Level 2 & 67 & 54 & 1 & 12 & 0.81 & 12 \\
\quad Level 3 & 11 & 3 & 3 & 5 & 0.27 & 5 \\
\bottomrule
\end{tabular}
\end{table}

Opus achieves 90\% overall accuracy with only 4 timeouts, compared to Sonnet's 78\% with 18 timeouts.
Both models show near-ceiling performance at Level~1 (95\%), but diverge sharply at higher difficulty: Opus maintains 93\% at Level~2 and 64\% at Level~3, while Sonnet drops to 81\% and 27\% respectively.
The 36-percentage-point gap at Level~3 ($p < 0.10$, Fisher's exact test) suggests that model scale disproportionately benefits deep constructive reasoning.

\paragraph{Performance by question type.}
Table~\ref{tab:pilot-by-type} reveals the example--counterexample gap for both models.

\begin{table}[ht]
\centering
\caption{Results by question type, showing the example--counterexample gap.}
\label{tab:pilot-by-type}
\begin{tabular}{lcccc}
\toprule
\textbf{Question Type} & \textbf{$n$} & \textbf{Opus Acc.} & \textbf{Sonnet Acc.} & \textbf{$\Delta$(O$-$S)} \\
\midrule
Example & 70 & 0.91 & 0.80 & $+$11pp \\
Counterexample & 23 & 0.83 & 0.65 & $+$18pp \\
Computation & 4 & 1.00 & 1.00 & 0 \\
Proof & 3 & 1.00 & 1.00 & 0 \\
\bottomrule
\end{tabular}
\end{table}

Both models show a gap between examples and counterexamples: approximately 9 percentage points for Opus (91.4\% vs.\ 82.6\%) and 15 for Sonnet (80.0\% vs.\ 65.2\%).
The larger Sonnet gap is partly driven by its higher timeout rate on counterexamples.
Notably, Opus's improvement over Sonnet is larger on counterexamples ($+$18pp) than on examples ($+$11pp), suggesting that model scale particularly benefits the ``concept stretching'' required for counterexample construction.

\paragraph{Performance by chapter.}
Table~\ref{tab:pilot-by-chapter} shows results across the four chapters.

\begin{table}[ht]
\centering
\caption{Results by chapter (Opus 4.6 primary, Sonnet 4.6 comparison).}
\label{tab:pilot-by-chapter}
\begin{tabular}{lccccc}
\toprule
\textbf{Chapter} & \textbf{$n$} & \textbf{Opus Acc.} & \textbf{Sonnet Acc.} & $\Delta$ & \textbf{Opus Partial} \\
\midrule
Cohomology & 15 & 1.00 & 0.93 & $+$7pp & 0\% \\
Basic Concepts & 47 & 0.91 & 0.79 & $+$13pp & 8.5\% \\
Moduli & 17 & 0.88 & 0.65 & $+$24pp & 0\% \\
Computations & 21 & 0.81 & 0.76 & $+$5pp & 9.5\% \\
\bottomrule
\end{tabular}
\end{table}

Opus achieves 100\% on Cohomology and near-ceiling on Basic Concepts (91\%).
The largest cross-model improvement is on Moduli ($+$24pp), where Opus's stronger reasoning about abstract constructions---deformation spaces, moduli functors---substantially outperforms Sonnet.
Computations shows the smallest gap ($+$5pp), suggesting that multi-step algebraic calculations remain challenging regardless of model scale.

\subsection{Case Studies}

We present three case studies from the Opus evaluation illustrating representative behaviors.

\paragraph{Case Study 1: Successful deep construction (\ecagid{0363}).}
This Level~3 problem asks to compute genus-0 Gromov--Witten invariants $\langle H^i \mid H^j \mid H^k \rangle$ for $\mathbb{P}^n$ using Schubert calculus on $\operatorname{Gr}(2,n+1)$ and derive the quantum cohomology ring.
Opus produced a detailed 5{,}443-character response correctly reducing the invariants to Schubert intersections via Segre classes, applying Pieri's rule, and deriving $QH^*(\mathbb{P}^n) \cong \mathbb{Z}[H,q]/(H^{n+1}-q)$.
Notably, Sonnet timed out on this same problem, illustrating the capability gap at the highest difficulty level.

\paragraph{Case Study 2: Property confusion (\ecagid{0174}, F2).}
This Level~2 problem asks to compute the class group of a specific scheme.
Opus produced a lengthy response demonstrating deep mathematical understanding, but the final class group calculation contained an error in identifying the correct divisor class relations, resulting in a partial credit score (0.5).
The failure mode is F2: the construction strategy was correct, but a subtle property confusion led to the wrong answer.

\paragraph{Case Study 3: Computational error (\ecagid{0256}, F3).}
This Level~2 problem asks to compute the Picard group of a specific variety.
Opus correctly identified the approach (intersection pairing analysis, non-projectivity argument) and arrived at the correct conclusion $\Pic(X) \cong \mathbb{G}_a$, but made an arithmetic error in the intermediate computation that happened to cancel out.
The scorer caught the error in the reasoning, assigning partial credit (0.5) despite the correct final answer.

\subsection{Cross-Model Comparison}

The comparison between Opus and Sonnet reveals three key patterns:

\begin{enumerate}[nosep]
  \item \textbf{Ceiling effects at Level~1.} Both models achieve 95\% at Level~1, suggesting that routine graduate-level constructions are within reach of mid-tier frontier models.
  \item \textbf{Scale benefits deep reasoning.} The accuracy gap widens monotonically with difficulty: 0pp at Level~1, 12pp at Level~2, 36pp at Level~3. This is driven both by fewer timeouts (Opus: 4 vs.\ Sonnet: 18) and by higher quality among completed responses.
  \item \textbf{Scale benefits counterexample construction.} The Opus--Sonnet gap is larger on counterexamples ($+$18pp) than on examples ($+$11pp), suggesting that the ``boundary awareness'' required for counterexample construction benefits more from model scale.
\end{enumerate}

\subsection{Comparison with Existing Benchmarks}

Direct comparison with other benchmarks is difficult due to differences in domain and evaluation methodology.
However, several reference points are informative:
\begin{itemize}[nosep]
  \item FrontierMath~\citep{glazer2024frontiermath}: frontier models solve $<$2\% of research-level problems, placing \ecagbench{} in an intermediate regime.
  \item MATH benchmark~\citep{hendrycks2021measuring}: frontier models achieve $>$90\%, confirming that competition mathematics is largely solved.
  \item PutnamBench~\citep{tsoukalas2024putnambench}: formal theorem proving success rates remain low for the hardest problems.
\end{itemize}

\ecagbench{} occupies an intermediate position.
The 90\% accuracy of the strongest model (Opus) on the full subset, combined with 64\% at Level~3, suggests that routine constructive algebraic geometry is accessible to frontier models but deep problems remain challenging.
The steep difficulty gradient (95\% $\to$ 93\% $\to$ 64\% for Opus) and cross-model comparison confirm that \ecagbench{} effectively spans the transition from solvable to challenging.
