\section{Related Work}
\label{sec:related}

\subsection{Mathematical Reasoning Benchmarks}

Table~\ref{tab:benchmark-comparison} provides a systematic comparison of \ecagbench{} with existing mathematical reasoning benchmarks.

\begin{table*}[t]
\centering
\caption{Comparison of mathematical reasoning benchmarks. ``Constructive?'' indicates whether problems require constructing mathematical objects (as opposed to producing proofs or unique answers). ``Counterex.'' indicates whether counterexample construction is explicitly tested.}
\label{tab:benchmark-comparison}
\small
\begin{tabular}{lrllccc}
\toprule
\textbf{Benchmark} & \textbf{Size} & \textbf{Domain} & \textbf{Verification} & \textbf{Construct.?} & \textbf{Counterex.?} & \textbf{Open?} \\
\midrule
MATH~\citep{hendrycks2021measuring} & 12{,}500 & Competition math & Exact match & \texttimes & \texttimes & \checkmark \\
AGIEval~\citep{zhong2023agieval} & 8{,}062 & Standardized tests & Multiple choice & \texttimes & \texttimes & \checkmark \\
miniF2F~\citep{zheng2022minif2f} & 488 & Olympiad math & Proof assistant & \texttimes & \texttimes & \checkmark \\
PutnamBench~\citep{tsoukalas2024putnambench} & 1{,}697 & Putnam competition & Lean/Isabelle/Coq & \texttimes & \texttimes & \checkmark \\
FrontierMath~\citep{glazer2024frontiermath} & 300+ & Research math & Custom scripts & Partial & \texttimes & \texttimes \\
GHOSTS~\citep{zheng2024ghosts} & 700+ & Graduate science & Rubric & \texttimes & \texttimes & \checkmark \\
AMBER~\citep{liu2025amber} & 500+ & Lean 4 math & Lean 4 & \checkmark & \texttimes & \checkmark \\
MathArena~\citep{matharena2025} & Varies & Recent competitions & Dual grading & \texttimes & \texttimes & \checkmark \\
\midrule
\ecagbench{} (ours) & 332 & Algebraic geometry & CAS + rubric & \checkmark & \checkmark & \checkmark \\
\bottomrule
\end{tabular}
\end{table*}

\paragraph{Competition and undergraduate mathematics.}
The MATH dataset~\citep{hendrycks2021measuring} evaluates competition-level mathematics across seven subjects, with problems requiring numerical or symbolic answers.
AGIEval~\citep{zhong2023agieval} includes math problems from standardized tests.
These benchmarks test \emph{problem-solving} ability but focus on high-school and undergraduate competition mathematics, where problems have unique answers and standard solution techniques.
\citet{frieder2023mathematical} systematically evaluated ChatGPT's mathematical capabilities, finding significant gaps in reasoning even on undergraduate-level problems.

\paragraph{Research-level mathematics.}
FrontierMath~\citep{glazer2024frontiermath} is the closest benchmark in spirit, comprising research-level mathematics problems with custom verification scripts.
However, FrontierMath spans all areas of mathematics and does not specifically target constructive reasoning; many problems still ask for unique numerical answers.
GHOSTS~\citep{zheng2024ghosts} evaluates graduate-level science tasks using rubric-based scoring, but covers mathematics only as one component of a broader scientific evaluation.
Minerva~\citep{lewkowycz2022minerva} demonstrated that language models pretrained on scientific text could achieve strong performance on quantitative reasoning, spurring interest in more challenging mathematical benchmarks.

\paragraph{Formal theorem proving.}
PutnamBench~\citep{tsoukalas2024putnambench} formalizes Putnam competition problems in Lean, Isabelle, and Coq, testing whether LLMs can produce machine-verifiable proofs.
AMBER~\citep{liu2025amber} combines construction and verification in Lean~4, requiring models to both state and prove lemmas.
miniF2F~\citep{zheng2022minif2f} provides a cross-system benchmark for Olympiad-level formal mathematics.
NaturalProver~\citep{welleck2022naturalprover} and ProofNet~\citep{azerbayev2023proofnet} explore the interface between informal and formal mathematical reasoning.
These benchmarks operate in the formal verification paradigm, where correctness is checked by a proof assistant.
\ecagbench{} differs by operating in the \emph{informal} mathematical setting, using CAS verification of object properties rather than formal proof checking.

\paragraph{Interactive and multi-turn evaluation.}
\citet{collins2024evaluating} propose evaluating mathematical capabilities through multi-turn interactions.
MathArena~\citep{matharena2025} uses recent competition problems with dual grading (automatic and human) to minimize contamination.
\ecagbench{} currently uses single-turn evaluation but is designed to support multi-turn interaction in future work.

\subsection{Domain-Specific Benchmarks}

While general mathematical benchmarks span many areas, there is growing recognition that domain-specific evaluation is essential for measuring deep mathematical competence.
AlphaGeometry~\citep{trinh2024alphageometry} demonstrated that combining neural language models with symbolic engines can solve Olympiad-level geometry problems, but geometry in the Euclidean sense is a substantially different domain from algebraic geometry over arbitrary fields and rings.
To our knowledge, \ecagbench{} is the \emph{first} benchmark specifically targeting algebraic geometry---a field that sits at the intersection of algebra, geometry, and topology, requiring fluency with abstract categorical constructions alongside concrete polynomial computations.

The domain specificity of \ecagbench{} is both a limitation and a strength.
On one hand, results do not directly generalize to other mathematical domains.
On the other hand, depth within a single domain allows us to calibrate difficulty levels against the actual pedagogical structure of algebraic geometry and to design verification strategies that exploit the domain's computational tools (SageMath~\citep{sagemath}, Macaulay2~\citep{macaulay2}, Singular~\citep{singular}).

\subsection{Constructive Reasoning and Counterexamples}

The distinction between constructive and verificational reasoning is well-established in mathematics~\citep{hartshorne1977algebraic} but has received limited attention in the LLM evaluation literature.
Lakatos~\citep{lakatos1976proofs} placed counterexamples at the center of mathematical methodology, arguing that mathematical concepts are refined through a dialectical process of conjecture, attempted proof, and refutation by counterexample.
In algebraic geometry, this process is especially pronounced: many fundamental results (e.g., Zariski's Main Theorem, cohomological criteria for ampleness) are accompanied by canonical counterexamples that delineate their precise scope.

\ecagbench{} explicitly tests both directions of this dialectic: 243 problems ask for examples (``construct an object satisfying\ldots'') and 66 ask for counterexamples (``construct an object violating\ldots'').
This dual structure is, to our knowledge, unique among mathematical benchmarks.
As we show in Section~\ref{sec:experiments}, the gap between model performance on examples versus counterexamples is substantial and informative.

\subsection{Contamination Considerations}

A persistent concern in LLM evaluation is \emph{data contamination}: test problems that appear in the training corpus.
\ecagbench{} has several properties that mitigate this risk:
\begin{enumerate}[nosep]
  \item The source material is a custom-written collection of examples, not drawn from standard problem sets, competition archives, or widely-distributed textbooks.
  \item Problems are embedded as structured annotations within the source material and extracted programmatically; the specific problem formulations do not exist as standalone documents on the web.
  \item The content enhancement process that produced the final problem statements was performed using the Anthropic API, generating unique formulations that are unlikely to match any training data verbatim.
  \item The source repository has been public for a limited period, reducing the probability that its content entered pre-training corpora for currently evaluated models.
\end{enumerate}
While no benchmark can guarantee zero contamination, these properties place \ecagbench{} at substantially lower contamination risk than benchmarks derived from competition archives or widely-used textbooks.
